Appendix: New and simplified manual controls for projection and slice tours

### 2022-07-12 paper/appendix.Rmd
Adjustment method 3 as algorithm

For now just sketching the idea:

- first step is to click on the axis display to change the contribution of one variable $m$
- capture that position and replace the corresponding row in the projection matrix
- for 2D projection this gives components $m_1$ along "x" direction and $m_2$ along $y$ direction
- next step: rotate basis such that direction of row m now corresponds to the first basis vector, this means we apply 2x2 rotation matrix to each row of the projection matrix, where the rotation angle is $tan(\theta) = m_2/m_1$ (that matrix can be written just in terms of that ratio, $x= m_2/m_1$, when $m_1<0$ I think we just need to translate $\theta \rightarrow \theta + \pi$)
- take the rotated basis, apply Gram-Schmidt, now the direction of $m$ will not change (but the length will change during normalization)
- rotated back to the original xy basis to update the plots

XXX need to check calculation of angle when $m_1 < 0$

>>>>

A very simple adjustment method works purely within the (hyper)plane, and is using the fact that the first direction will not be rotated when applying Gram-Schmidt. The following steps are needed:

1. Provide $A$, and $m$.
2. Change values in row $m$, giving $A^*$.
3. Rotate the basis within the projection (hyper-)plane, such that the first direction is aligned with the direction of $m$ in $A^*$, giving $\hat A^*$. In a 2D projection this means finding the angle $\theta$ by which the basis should be rotated.
4. Normalise the first direction in $\hat A^*$ while keeping the contribution in row $m$ fixed, giving $\hat A^{*0}$.
5. Orthonormalise $\hat A^{*0}$ using Gram-Schmidt, giving $\hat A^{**}$. By construction the contribution of $m$ will remain fixed.
6. Rotate $\hat A^{**}$ back into the original basis to obtain $A^{**}$.

Diagram needed?

### 2022-07-12 paper/appendix.tex
Adjustment method 3 as algorithm

For now just sketching the idea:

\begin{itemize}
\tightlist
\item
  first step is to click on the axis display to change the contribution
  of one variable \(m\)
\item
  capture that position and replace the corresponding row in the
  projection matrix
\item
  for 2D projection this gives components \(m_1\) along ``x'' direction
  and \(m_2\) along \(y\) direction
\item
  next step: rotate basis such that direction of row m now corresponds
  to the first basis vector, this means we apply 2x2 rotation matrix to
  each row of the projection matrix, where the rotation angle is
  \(tan(\theta) = m_2/m_1\) (that matrix can be written just in terms of
  that ratio, \(x= m_2/m_1\), when \(m_1<0\) I think we just need to
  translate \(\theta \rightarrow \theta + \pi\))
\item
  take the rotated basis, apply Gram-Schmidt, now the direction of \(m\)
  will not change (but the length will change during normalization)
\item
  rotated back to the original xy basis to update the plots
\end{itemize}

XXX need to check calculation of angle when \(m_1 < 0\)

>>>>

A very simple adjustment method works purely within the (hyper)plane,
and is using the fact that the first direction will not be rotated when
applying Gram-Schmidt. The following steps are needed:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Provide \(A\), and \(m\).
\item
  Change values in row \(m\), giving \(A^*\).
\item
  Rotate the basis within the projection (hyper-)plane, such that the
  first direction is aligned with the direction of \(m\) in \(A^*\),
  giving \(\hat A^*\). In a 2D projection this means finding the angle
  \(\theta\) by which the basis should be rotated.
\item
  Normalise the first direction in \(\hat A^*\) while keeping the
  contribution in row \(m\) fixed, giving \(\hat A^{*0}\).
\item
  Orthonormalise \(\hat A^{*0}\) using Gram-Schmidt, giving
  \(\hat A^{**}\). By construction the contribution of \(m\) will remain
  fixed.
\item
  Rotate \(\hat A^{**}\) back into the original basis to obtain
  \(A^{**}\).
\end{enumerate}

Diagram needed?
